---
title: "Data Transformation"
author: "Emile Vlahos"
date: "18/09/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5.1 Introduction 

### 5.1.1 Prerequisites

```{r}
library(tidyverse)
library(nycflights13)
```

# 5.1.2 nycFlights13

```{r}
flights
```

### 5.1.3 dplyr

verbs

filter 
arrange
select
mutate
summarize

verb(database, whatToDo?)

# 5.2 Filter rows with filter()

Flights that occurred on Jan 1st

```{r}
filter(flights, month == 1, day == 1)
```

```{r}
(jan1 <- filter(flights, month == 1, day == 1))
```

```{r}
(dec25 <- filter(flights, month == 12, day == 25))
```

### 5.2.1 Comparisons

==, >, >=, <, <=, !

```{r}
filter(flights, month == 1)
```

```{r}
cos(pi/2) == 0
```

```{r}
near(cos(pi/2), 0)
```

### 5.2.2 Logical operators

or ------> |
and ------> &
not ------> !

Flights that departed in November **or** December  

```{r}
filter(flights, month == 11 | month == 12)
```

may, jun, july, aug

```{r}
filter(flights, month == 5 | month == 6 | month == 7 | month == 8)
```

```{r}
filter(flights, month %in% c(5,6,7,8) )
```

Flights that departed in November **and** arrived in December

```{r}
# tbd
```

Flights that weren't delayed (on departure or arrival) by more than two hours

```{r}
filter(flights, dep_delay <= 120 & arr_delay <= 120)
```

##### 5.2.3 Missing Values

One important feature of R that make comparisons tricky are missing values, or NAs. NA represents an unkown value so missing values are contagious: almost any operation invloving an unkown value will also be unkown

```{r}
100 == NA
2 > NA
x <- NA
is.na(x)
```

How do I get rid of NA's on my dataset?

```{r}
filter(flights, is.na(dep_time))
filter(flights, !is.na(dep_time))
withoutNA <- filter(flights, !is.na(dep_time))
withoutNA

```

##### 5.2.4 Exercises

> 1. Find all flights that

Had an arrival delay of two or more hours

```{r}
filter(flights, arr_delay >= 120)
```

Flew to Houston (IAH or HOU)

```{r}
```

Departed in summer (July, August, and September)

```{r}

```

Arrived more than two hours late, but didn't leave late
 
```{r}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```

Were delayed by at least an hour, but made up over 30 minutes in flight


```{r}
select(flights, year, month, day)
select(flights, year:day)

select(flights, -(year:day))
select(flights, dep_time:time_hour)
```


### 5.6 Grouped Summaries with su,marise()

The last key verb is sumarize(). It collapses a data frame to a single row:

```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_dest <- group_by(flights, dest)
```

```{r}
delays <- summarize(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
  )

delays
```

```{r}
ggplot(data = delays) + 
  geom_point(mapping = aes(x = dist, y = delay, size = count))
```

```{r}
delays <- filter(delays, count > 20, dest != 'HNL')
```


```{r}
ggplot(data = delays) + 
  geom_point(mapping = aes(x = dist, y = delay, size = count))
```

```{r}
ggplot(data = delays, mapping = aes(x = dist, y = delay)) + 
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

```{r}
delays <- flights %>% 
  group_by(dest) %>%
  summarize(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>%
  filter(count > 20, dest != 'HNL')

ggplot(data = delays, mapping = aes(x = dist, y = delay)) + 
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
  
```


##### 5.6.2 Missing values

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarize(dep_delay = mean(dep_delay))

```





